---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)

p_load(stringr, dplyr, rethinking, ggplot2, gridExtra)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions


Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r setup, include=FALSE}
 
#Grid approximation code - considering only a finite grid of parameter values to achieve approximation og the posterior

#Ric = 3/6

dens = 100
#define grid
p_grid = seq(from=0, to = 1, length.out = dens)

#define prior
prior = rep(1,100)

#compute/define likelihood at eah value in grid
likelihood = dbinom(3, size=6, prob=p_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Ric_data= data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Ric_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")



#Implement quadratic approximation
Ric = map(
alist(
w ~ dbinom(6,p) , # binomial likelihood
p ~ dunif(0,1) # uniform prior
) ,
data=list(w=3) )
# display summary of quadratic approximation
precis( Ric )


#How much posterior falls below 50 %
sum(posterior[p_grid <0.5])
# 50 % chance on both sides. Bigger chance that he knows less and more.

```


2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.

```{r setup, include=FALSE}

#Kri = 2/2

dens = 100
#define grid
k_grid = seq(from=0, to = 1, length.out = dens)

#define prior
prior = rep(1,100)

#compute/define likelihood at eah value in grid
likelihood = dbinom(2, size=2, prob=k_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Kri_data= data.frame(grid=k_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Kri_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")


#Implement quadratic approximation
Kri = map(
alist(
w ~ dbinom(2,p) , # binomial likelihood
p ~ dunif(0,1) # uniform prior
) ,
data=list(w=2) )
# display summary of quadratic approximation
precis( Kri )


#How much posterior falls below 50 %
sum(posterior[k_grid <0.5])
# 12 % chance on both sides. So chances are high that he has absolute cogsci superpowers, but also that a man knows nothing. 



#Josh = 160/198

dens = 100
#define grid
J_grid = seq(from=0, to = 1, length.out = dens)

#define prior
prior = rep(1,100)

#compute/define likelihood at eah value in grid
likelihood = dbinom(160, size=198, prob=J_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Jos_data= data.frame(grid=J_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Jos_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")


#Implement quadratic approximation
Jos = map(
alist(
w ~ dbinom(198,p) , # binomial likelihood
p ~ dunif(0,1) # uniform prior
) ,
data=list(w=160) )
# display summary of quadratic approximation
precis( Jos )


#How much posterior falls below 50 %
sum(posterior[J_grid <0.5])
# VERY low chance that Josh has absolute superpowers, but also that he knows nothing. Very stable teacher. aPeak past 50. 



#Mik = 66/132

dens = 100
#define grid
M_grid = seq(from=0, to = 1, length.out = dens)

#define prior
prior = rep(1,100)

#compute/define likelihood at eah value in grid
likelihood = dbinom(66, size=132, prob=M_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Mik_data= data.frame(grid=M_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Mik_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")


#Implement quadratic approximation
Mik = map(
alist(
w ~ dbinom(6,p) , # binomial likelihood
p ~ dunif(0,1) # uniform prior
) ,
data=list(w=3) )
# display summary of quadratic approximation
precis( Mik )


#How much posterior falls below 50 %
sum(posterior[M_grid <0.5])
# 50 % chance on both sides. Chances of falling above or below 50 is much less than Ric, so very stable at 50 % whereas Ric cn be VERY good but can also be VERY bad where Mik stays 50 %.





#more skeptic = lower the priorm or assume a normal distribution

```


3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

```{r setup, include=FALSE}

#Kristian

dens = 100
#define grid
k_grid = seq(from=0, to = 1, length.out = dens)

#define prior, assume normality and change mean to 0.8 and sd to 0.2
prior = dnorm(k_grid, mean = 0.8, sd = 0.2)

#compute/define likelihood at eah value in grid
likelihood = dbinom(2, size=2, prob=k_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Kri_data= data.frame(grid=k_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Kri_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")




#Josh = 160/198

dens = 100
#define grid
J_grid = seq(from=0, to = 1, length.out = dens)

#define prior
prior = dnorm(J_grid, mean = 0.8, sd = 0.2)

#compute/define likelihood at eah value in grid
likelihood = dbinom(160, size=198, prob=J_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Jos_data= data.frame(grid=J_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Jos_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")



#Mik = 66/132

dens = 100
#define grid
M_grid = seq(from=0, to = 1, length.out = dens)

#define prior
prior = dnorm(M_grid, mean = 0.8, sd = 0.2)

#compute/define likelihood at eah value in grid
likelihood = dbinom(66, size=132, prob=M_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Mik_data= data.frame(grid=M_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Mik_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")




#Ric = 3/6

dens = 100
#define grid
p_grid = seq(from=0, to = 1, length.out = dens)

#define prior
prior = dnorm(p_grid, mean = 0.8, sd = 0.2)

#compute/define likelihood at eah value in grid
likelihood = dbinom(3, size=6, prob=p_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Ric_data= data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Ric_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")



```

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r setup, include=FALSE}

#Kristian

dens = 100
#define grid
k_grid = seq(from=0, to = 1, length.out = dens)

#define prior, assume normality and change mean to 0.8 and sd to 0.2
prior = dnorm(k_grid, mean = 0.8, sd = 0.2)

#compute/define likelihood at eah value in grid
likelihood = dbinom(200, size=200, prob=k_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Kri_data= data.frame(grid=k_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Kri_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")


#Kristian seems to be extremely knowledgeable unless we assume he cheated



#Josh = 160/198

dens = 100
#define grid
J_grid = seq(from=0, to = 1, length.out = dens)

#define prior
prior = dnorm(J_grid, mean = 0.8, sd = 0.2)

#compute/define likelihood at eah value in grid
likelihood = dbinom(16000, size=19800, prob=J_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Jos_data= data.frame(grid=J_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Jos_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")

#Josh has even more certainty in his abilities (or lack of in comparison with kristian)



#Mik = 66/132

dens = 100
#define grid
M_grid = seq(from=0, to = 1, length.out = dens)

#define prior
prior = dnorm(M_grid, mean = 0.8, sd = 0.2)

#compute/define likelihood at eah value in grid
likelihood = dbinom(6600, size=13200, prob=M_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Mik_data= data.frame(grid=M_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Mik_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")

#same as josh, maybe a slight bit worse


#Ric = 3/6

dens = 100
#define grid
p_grid = seq(from=0, to = 1, length.out = dens)

#define prior
prior = dnorm(p_grid, mean = 0.8, sd = 0.2)

#compute/define likelihood at eah value in grid
likelihood = dbinom(300, size=600, prob=p_grid)

#compute the product og the prior
unstd_posterior = likelihood * prior

#standardize posterior so it sums to 1
posterior = unstd_posterior/sum(unstd_posterior)


#replicate different priors
#prior = ifelse(grid <0.5,0,1)
#prior = exp(-5*abs(grid-0.5))


Ric_data= data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(Ric_data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("probability of water")+ ylab("posterior probability")

#Riccardo shows a very average knowledge, but much below our prior (what we expect of our teacher)


#Why?
#Collecting more data makes you more 



#seeing the plots together

gridExtra::grid.arrange(Kri_data, Jos_data, Mik_data, Ric_data)
#Doesn't work?

```

5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

```{r setup, include=FALSE}

#I would lower the prior to chance level, since teachers knowledge should be above chance to be recognized as knowledgeable. 



```


6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



